---
title: "Newsgroup Data"
author: "Carson Sievert"
date: "May 28, 2014"
output: 
  html_document:
    self_contained: false
    theme: united
    highlight: tango
---

[LDAvis](https://github.com/cpsievert/LDAvis/) comes prepackaged with some data sets to help quickly demonstrate how to use it. This document visualizes a topic model fit to Newsgroup documents^[See http://qwone.com/~jason/20Newsgroups] created with **LDAvis** and **knitr** ([see here]() for source code). There are four essential components to this data structure:

```{r data, message = FALSE}
library(LDAvis)
data("Newsgroupdata", package = "LDAvis")
str(Newsgroupdata)
```

The first element of this list - "phi" - is a matrix and is one of the main outputs from an [LDA topic model](http://en.wikipedia.org/wiki/Latent_Dirichlet_allocation). Each column of this $\phi$ matrix defines a probability mass function over terms for a given topic. Consequently, the columns of the $\phi$ matrix should sum to 1 (roughly).

```{r sum}
with(Newsgroupdata, colSums(phi))
```

In this case, we've included terms in the rownames of the $\phi$ matrix which also defines the vocabulary (i.e., the set of all terms). 

```{r vocab}
all(with(Newsgroupdata, rownames(phi) == vocab))
```

The `term.frequency` is simply the number of times each term appears in the entire corpus where `topic.proportion` contains the percentage of tokens that derive from each topic. **LDAvis** comes equipped with a `check.inputs` function which conducts some basic checks on each of these elements.

```{r inputs}
# Check the inputs and sort topics by frequency:
z <- with(Newsgroupdata, check.inputs(K = 50, W = 22524, 
                phi, term.frequency, vocab, topic.proportion))
```

Most importantly, `check.inputs` will reorder columns of the $\phi$ matrix based on `topic.proportion` so that "first topic" represents the most frequent topic.

```{r reorder}
with(Newsgroupdata, order(topic.proportion, decreasing = TRUE))
with(z, colnames(phi))
```

For this reason, it's a good idea to relabel the column names of $\phi$

```{r relabel}
colnames(z$phi) <- seq_len(z$K)
``` 

At this point, we have an option to create a [shiny](http://shiny.rstudio.com/) based visualization with `runVis` or we can `createJSON` to derive a JSON object that will feed a standalone webpage. Although the shiny based visualization has a few more controls, the standalone page allows us to browse relevant terms for different topics while preserving [object constancy](http://bost.ocks.org/mike/constancy/) (try clicking on one of the circles below then decreasing the value of $\lambda$). `createJSON` fosters this approach by recomputing the top 30 most relevant terms for each topic (over a grid of values for $\lambda$).^[See [here](http://nlp.stanford.edu/events/illvi2014/papers/sievert-illvi2014.pdf) for the definition and discussion of relevance.]

```{r json, results = 'hide', cache = TRUE}
json <- with(z, createJSON(K, phi, term.frequency, vocab,
                            topic.proportion, n.terms=30, dist.measure = "JS"))
```

Now that we have `json`, we can use `serVis` function to spit out files required to view the interactive webpage. One can even use this function to upload files as [a gist](https://gist.github.com/cpsievert/70ca32cd3f4af8fe4cd4) which can then be viewed via [bl.ocks.org](http://bl.ocks.org/cpsievert/raw/70ca32cd3f4af8fe4cd4/).

```{r serVis}
serVis(json, out.dir = "newsgroup", open.browser = FALSE)
```

Now, to embed the resulting [/newsgroup/index.html](/newsgroup/index.html) webpage within this page, we can make use of the HTML `iframe`s.

<iframe src = "newsgroup/index.html" width=1200 height=700></iframe>

The `createJSON` function also takes an argument allowing us to change how the distance between topics is measured. Notice how using symmetric Kullback-Leibler (as opposed to Jensen-Shannon) divergence alters the locations of points on the left-hand side of the visualization.

```{r json2, results = 'hide', cache = TRUE}
# Takes 1-2 minutes to set up the data:
json2 <- with(z, createJSON(K, phi, term.frequency, vocab,
                            topic.proportion, n.terms = 30, dist.measure = "KL"))
```

```{r serVis2}
serVis(json2, out.dir = "newsgroup2", open.browser = FALSE)
```

<iframe src = "newsgroup2/index.html" width=1200 height=700></iframe>


