---
title: "Using LDAvis with shiny & rmarkdown"
author: "Carson Sievert"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

This document details how to make **LDAvis** react to [**shiny**](http://cran.r-project.org/web/packages/shiny/index.html) inputs. In theory, any `createJSON()` argument *could* be dynamically altered in a [shiny app](http://shiny.rstudio.com/) or [interactive document](http://rmarkdown.rstudio.com/authoring_shiny.html). In practice, this might not be such a great idea since `createJSON()` can take a few seconds to run (note in this example it's about 5 seconds).

Anyway, here is one way to alter the number of terms shown using a `shiny::sliderInput()`.

```{r}
library(shiny)
sliderInput("nTerms", "Number of terms to display", min = 20, max = 40, value = 30)
```

By wrapping a call to `createJSON()` with `renderVis()`, the output will update whenever the value of `input$nTerms` changes.

```{r}
library(LDAvis)
data(TwentyNewsgroups, package = "LDAvis")
renderVis({
  with(TwentyNewsgroups, 
       createJSON(phi, theta, doc.length, vocab, term.frequency, 
                  R = input$nTerms))
})
```